\documentclass[a4paper,12pt]{article}
\usepackage{pf2}
\usepackage{amssymb}
\usepackage{ucs}
\usepackage[utf8x]{inputenc}
\begin{document}

\section{theorem}
$f:X→Y$ has inverse iff it is bijective.

\subsection{lemma}
For two functions $f:X\rightarrow Y, g:Y→X$ whenever $gf=e_X$, then f is injective (into, no one image has more than one pre-image) and g is surjective (onto, all of $Y$ is mapped onto $X$).

%% huh, why is there indentation?

($e_M$ is an identity automorphism on set $M$.)

\subsection{proof of the lemma}

Consider an element $x$ in $X$ such that $g(f(x))=e_X(x)=x$. For any two $x, x' \in X$ if $f(x) = f(x') = y$, then $g(y) = x = x'$, thus $x=x'$.

\vspace{2ex}
Definition: a function $f:X→Y$ is injective, or into, if $\forall x, x' \in X   : f(x)=f(x') \Rightarrow x=x' $. 

%HWÆT how does it follow that $g$ is surjective?

That is to say, there are no such element of range,$y \in Y$, that has two different pre-images $x, x' \in X: y=f(x)=f(x')$.

\vspace{2ex}
Definition: a function $g:X→Y$ is surjective, or onto, if $\forall y \in Y \thinspace  | \exists x \in X: g(y)=x$. 

That is to say, all elements of the range have pre-images.
\subsection{proof of the theorem}

$\textsc{Bijective}(f) \Leftarrow ∃ f^{-1}:$\\

Bijective means that $ ∀ x \in X ∃! y \in Y: f(x)=y$ Since y is unique, we can begin to construct $f^{-1}: Y → X$ with $(y,x)$. There is such a pair of every $x\in X$ and every $x \in Y$ and all such pairs are unique by bijectivity.\\

This is a mess, not a proof.\\
What do I need to do so that the theorem is properly shown to be true? Unfold the definitions and show that the pieces align accoring to the rules of derivation.
So.

 $\textsc{Bijective}(f)\triangleq\textsc{Injective}(f)∧\textsc{Surjective}(f)$\\ 
Expanding further,\\
 $\textsc{Bijective}(f)= (∀y ∃x: y=f(x))∧(∀x, x': f(x)=f(x')\Rightarrow x=x')$\\ 
\\
$ ∃ f^{-1} \Leftarrow \textsc{Bijective}(f):$\\
\\

So far so good. Now, what does it mean for a function to have inverse?
It means that the function is one-to-one \textit{and} that it covers the whole range.

Injectivity and surjectivity are not symmetric with respect to domain and range: function is always considered defined on all of the domain, but we specifically call /injective/ those functions for which all of their range is used. Any non-injective function can be narrowed down to the used part of its range to make an injective function. More formally: $∀ f, f:X→Y, ∃ f':X→Y' where Y' =\{y \in Y: ∃x \in X: f(x) = y\}.$
\\

In contrast, non-surjective functions can't be narrowed... no, wait, they can! their /domain/ can be narrowed so that for any surjective $f : X→Y$ we can construct $f':X'→Y$ where $X'=X\{z \in X, ∃ x \in X, x \neq z, f(x)=f(z)$. Nope that formalization excludes all non-unique pre-images, but I want to retain one. Actually, the formalization works: what remains /is/ a surjective function, a minimal one in a large family. Obviously, we often have additional properties we want our surjective narrowing to have, say, analiticity. This seems to be amenable to formalization via an ordering on $X$, where we simply retain the smallest (or the largest) element, and throw away the rest.
\\

So surjectivity says something about the \textit{domain}: none of its elements have matching images, the function is one-to-one

Injectivity is a statement about the range: the whole of the range is covered.

When both are true, the function is nicely one-to-one and there are no holes in its range.

\pagebreak{}
\section{Lamport-style structured proof} 
% HWÆT: where does the statement of the theorem go? 
% WOOT! Right before the first step of the proof, with title in bold. 
\textsc{Theorem}: When a composition of two functions is an identity, the functions are inverses of each other and are bijective. 
% Bad style! what are g and f? what is X and id_X?
% HWÆT: LaTeX, how to manipulate vertical space? Also, how much space should there be? What is the proper typographical reasoning here?

What follows as an attempt to prove the previous statement in line with Leslie Lamport's article "How to write a 21-century proof" using his \texttt{pf2.sty}.
\\
It suffices to prove that bijectivity is equivalent to the existence of inverse, and ... Dammit, this is trivial but unclear how to prove.

\begin{proof}
  \step{theorem}{$\textsc{Bijective}(f) \Leftrightarrow ∃ f^{-1}$}
  \begin{proof}
    \step{lemma}{$g \circ f = id_X \Rightarrow \textsc{Injective}(f) ∧ \textsc{Surjective}(g)$}
    \begin{proof}
          \step{from id surjectivity}{$g \circ f = id_X \Rightarrow \textsc{Surjective}(g)$}
          \begin{proof}
%            \pf by ($g \circ f = id_X \Rightarrow ∀x \in X : x = id_X(x) = (gf)x = g(fx)$) and definition of Surjectivity.~\qed
            \pf by ($g \circ f = id_X \Rightarrow ∀ x ∃ f(x) \in Y : x = id_X(x) = g \circ f (x) = g(f(x))$) and definition of Surjectivity.~\qed
% this step may be written out more clearly.
          \end{proof}
          \step{from id injectivity}{ $g \circ f = id_X \Rightarrow \textsc{Injective} (f)$}
          \begin{proof}
            \pf by definition of Injectivity.~\qed
          \end{proof}
          \qedstep
        \end{proof}
 % \step{def bijectivity}{$\textsc{Bijective}(f)\triangleq\textsc{Injective}(f)∧\textsc{Surjective}(f)$}
 %  \pf\Paragraph By definition. ~\qued
 %  \step{unwind \textit{further}}
 %  {$\textsc{Bijective}(f) = 
 %     (∀y ∃x: y=f(x))
 %    ∧(∀x, x' : f(x)
 %    =f(x') \Rightarrow x=x')$}


\end{proof}
\end{proof}
  

% RESEARCH: Zermelo-Frænkel set notation. Should I write  $\forall x \in X$ or what? How do I write "such that"?
% gripe: make $ insert a pair of $s, similar to various brackets.
% how to make new emacs commands? I might use one to jump to this commented out section, or even to keep it in a separate buffer.
% 
% HWÆT: how to map latex command onto article structure? sections, theorems, lemmas, proofs, do all of these live on the \subsection level, or even below?
% HWÆT: I need slurping practice. And slurping cheatsheet!

% RESEARCH: how to make latex accept unicode arrows in the input file?
% on a related note, how do I use Russian (or Ethiopian, for that matter)
% while retaining English as a document-level language?
% What's the difference between babel and the other options (polyglossia? what else?)

\end{document}
